Make_mash_U_hyp

### 2020-12-03 R/Analysis_v0.5_mash_GxE.Rmd
Move R scripts not used in current manuscript and rename those that are

@@

Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?
```{r}
log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
    as.data.frame() %>%
    rownames_to_column(var = "value") %>%
    mutate(value = as.integer(.data$value)) %>%
    as_tibble() %>%
    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
    dplyr::rename(log10BayesFactor = .data$V1) %>%
    dplyr::select(-.data$value)

>>>>

# TRIAL script
# -------------------------
NB: Trial of generating covariance matrices for combinations of environmental phenotypes from the U_ed from single environmental phenotypes.

@@

# --- New Mashhattan --
Mashhattan where points are colored by which covariance matrix they load most strongly on to. Do different regions of the genome have distinct patterns, or are they all jumbled up?
```{r}
log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
    as.data.frame() %>%
    rownames_to_column(var = "value") %>%
    mutate(value = as.integer(.data$value)) %>%
    as_tibble() %>%
    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
    dplyr::rename(log10BayesFactor = .data$V1) %>%
    dplyr::select(-.data$value)

### 2020-12-03 R/Plots_v0.2_manuscript_figures.Rmd
Move R scripts not used in current manuscript and rename those that are

  for(i in 6:9){
    for(j in seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

@@

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_", 
                                                  "functions_of_weather_v2.csv")))
}

>>>>

for(p in 2){#seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE):vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

@@

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_", 
                                                  "functions_of_weather_v3.csv")))
}

### 2020-12-03 R/Plots_v0.3_G_BLUP.Rmd
Move R scripts not used in current manuscript and rename those that are

```{r}

@@

Do the QTL overlap any of the genes in ftanno?
```{r}
ft_annos %>%
  filter(seqid %in% qtldf$CHR[1] & between(Pos_Mb, qtldf$Pos_Mb_lo[1], qtldf$Pos_Mb_hi[1]))

qtldf
i=24
ft_annos %>%
  filter(seqid %in% qtldf$CHR[i] & between(Pos_Mb, qtldf$Pos_Mb_lo[i], qtldf$Pos_Mb_hi[i])) %>%
  select(OsGene, AtGene, everything())
```

>>>>

# Total # 20kb regions
Midwest: 8824339 SNPs       52797 20kb regions
Gulf: 10320201 SNPs         53025 20kb regions
Gulf_and_Midwest: 12083955  53086 20kb regions
53240 20kb regions total
```{r}
phe_v <- c("dyln_fl50", "cgdd_12c_gr2fl", "FL50")
site_v <- c("eight_sites", "tx_sites", "north_sites")
i=1
j=1
k=1

  gwas_files <- list.files(path = paste0("./", site_v[j]), pattern = paste0("GWAS_datatable_", phe_v[i]))

  gwas_files

  gwas_20kb <- tibble()
subpop_v <- c("Midwest", "Gulf", "Gulf_and_Midwest")

for(k in seq_along(gwas_files)){
  gwas1 <- readRDS(file.path(site_v[j], gwas_files[k]))
gwas1 <- gwas1 %>%
  mutate(Pos_Mb = round(POS/10000)/100,
         POS_bin = round(Pos_Mb*100/2)) %>%
  group_by(CHR, POS_bin) %>% 
  tally() %>%
  mutate(subpop = subpop_v[k])
gwas_20kb <- bind_rows(gwas_20kb, gwas1)
}
```

# Overlaps phe across subpops

Associations did or did not overlap significantly in the same phenotypes among subpopulations? What about in the same subpop between sets of sites? This is GxE at the continental scale, point out.
In both Gulf & Midwest:
All  dyln_fl50  Chr01N	229	2	   1 region 1/463     NS 0.08496942
All cgdd_12c_gr2fl  0 regions   0/2
All FL50  0 regions   0/18
North  dyln_fl50 39 regions  39/2530      sig 0.00688
North cgdd_12c_gr2fl  0 regions  0/59
North  FL50    6 regions 5Nx2 6Kx2 8N 9K   6/1097     NS 0.141
Texas  dyln_fl50         0 regions  0/232
Texas  cgdd_12c_gr2fl    0 regions  0/190
Texas  FL50             0 regions   0/64
```{r}
sitev <- "eight_sites"
phev <- "dyln_fl50"
popv <- list(Gulf = 53025, Midwest = 52797, Gulf_and_Midwest = 53086)
i=2

AandB <- gwas_fdr %>%
  filter(subpop %in% names(popv)[c(1,2)] & phe %in% phev & sites %in% sitev) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n()) %>%
    filter(count > 1)
inB <- gwas_fdr %>%
  filter(subpop %in% names(popv)[1] & phe %in% phev & sites %in% sitev) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n()) 
notinB <- popv[[i]] - nrow(inB)
inA <- gwas_fdr %>%
  filter(subpop %in% names(popv)[2] & phe %in% phev & sites %in% sitev) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n())

phyper(nrow(AandB), nrow(inB), notinB, nrow(inA), lower.tail = FALSE)
```

Eight & Texas:
In Gulf
dyln_fl50  22/274
FL50
cgdd_12c_gr2fl 0/2

## Hypergeometric test for significance of overlaps
phyper(10, 50, 200 - 50, 25, lower.tail = FALSE)
phyper(A&B, inB, notinB, inA, lower.tail = FALSE)
```{r}
A <- "north_sites"
B <- "tx_sites"
phev <- "cgdd_12c_gr2fl"
popv <- list(Gulf = 53025, Midwest = 52797, Gulf_and_Midwest = 53086)
i=2

AandB <- gwas_fdr %>%
  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(A,B)) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n()) %>%
    filter(count > 1)
inB <- gwas_fdr %>%
  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(B)) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n()) 
notinB <- popv[[i]] - nrow(inB)
inA <- gwas_fdr %>%
  filter(subpop %in% names(popv)[i] & phe %in% phev & sites %in% c(A)) %>%
    group_by(sites, subpop, phe, CHR, POS_bin) %>%
    tally() %>%
    group_by(CHR, POS_bin) %>%
    summarise(count = n())

phyper(nrow(AandB), nrow(inB), notinB, nrow(inA), lower.tail = FALSE)
8/232
```
tx & north dyln Gulf 0.0078 (8 overlaps)
eight & north cgdd Gulf 1.13e-06
eight & north dyln Gulf 5.36e-31 (25 overlaps)
eight & tx dyln Gulf 2.2e-38 (22 overlaps)

eight & north dyln Midwest 4.42e-242
Very few overlaps between site sets for the Midwest subpopulation.
Lots of overlaps & sig for Gulf_and_Midwest, if we trust this.

@@

Do the QTL overlap any of the genes in ftanno?
```{r}
ft_annos %>%
  filter(seqid %in% gtldf3$CHR[1] & between(Pos_Mb, gtldf3$Pos_Mb_lo[1], gtldf3$Pos_Mb_hi[1]))

gtldf3 <- qtldf %>% filter(lod > 30 & lodcolumn == "dyln_fl50")
i=4
ft_annos %>%
  filter(seqid %in% gtldf3$CHR[i] & between(Pos_Mb, gtldf3$Pos_Mb_lo[i], gtldf3$Pos_Mb_hi[i])) %>%
  select(OsGene, AtGene, everything()) %>%
  left_join(osgene_ft_key, by = c("OsGene" = "Symbol"))
```
## start here QTL candidate genes
Chr02N	57.23	58.81	   SIP1|OsSUF4	SUF4
SIP1 participates in regulation of flowering time in rice by recruiting OsTrx1 to Ehd1.

Chr04K	4.23	5.50     6 genes including Hd3a	FT
	Hd3a, a rice ortholog of the Arabidopsis FT gene, promotes transition to flowering downstream of Hd1 under short-day conditions
	Phytochrome dependent quantitative control of Hd3a transcription is the basis of the night break effect in rice flowering
	Variations in Hd1 proteins, Hd3a promoters, and Ehd1 expression levels contribute to diversity of flowering time in cultivated rice

Chr04K	9.92	13.93    3 genes including Hd3a	FT ( a different copy)
Hd3a, a rice ortholog of the Arabidopsis FT gene, promotes transition to flowering downstream of Hd1 under short-day conditions
	Phytochrome dependent quantitative control of Hd3a transcription is the basis of the night break effect in rice flowering
	OsMFT1 increases spikelets per panicle and delays heading date in rice by suppressing Ehd1, FZP and SEPALLATA-like genes.

Chr05N	2.04	3.44     5 genes including FT and OsFTL1	FT &&&&& OSMADS3:
Genetic Enhancer Analysis Reveals that FLORAL ORGAN NUMBER2 and OsMADS3 Cooperatively Regulate Maintenance and Determinacy of the Flower Meristem in Rice.

Chr05N	64.27	65.98	  3 genes including OsFTL10 and OsABF1|OsABI5|OREB1|OsbZIP10: 	A Drought-inducible bZIP Transcription Factor OsABF1 Delays Reproductive Timing in Rice.

### 2020-12-03 R/Pre_manuscript/1_bigsnpr_gwas_prep.Rmd
Move R scripts not used in current manuscript and rename those that are

>>>>

## Visualize mash effects

```{r}
mash_output <- readRDS("FL50/Strong_Effects29475SNPs.rds")
```

```{r}
i=10
get_significant_results(mash_output)[i]
plot1 <- switchgrassGWAS::mash_plot_effects(m = mash_output, n=i)

xlabs <- str_sub(plot1$effect_df$value, start = 11, end = -6)
plot1$effect_df$SITE <- xlabs

#plot1$effect_df$SITE <- factor(plot1$effect_df$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
plot1$effect_df$SITE  <- factor(plot1$effect_df$SITE , levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))

plot1$ggobject + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1)) + 
  scale_x_discrete(labels = xlabs)

ggplot(data = plot1$effect_df) +
    geom_point(mapping = aes(x = .data$SITE, y = .data$mn)) +
    geom_errorbar(mapping = aes(ymin = .data$mn - .data$se,
                                ymax = .data$mn + .data$se,
                                x = .data$SITE), width = 0.3) +
    geom_hline(yintercept = 0, lty = 2) +
    labs(x = "Conditions", y = "Effect Size") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

save_plot(paste0(names(get_significant_results(mash_output)[i]), "_Effect_Plot.png"), plot = last_plot(), base_height = 4, base_asp = 0.8)

### 2020-12-03 R/Pre_manuscript/D2F_breakdown_by_site.Rmd
Move R scripts not used in current manuscript and rename those that are

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
```

>>>>

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
```

### 2020-12-03 R/Pre_manuscript/Draft_figures_2020-06-02.Rmd
Move R scripts not used in current manuscript and rename those that are

#
```{r}
metadata <- read_rds("../data/metadata.rds")
subpops <- read_rds("../data/subpop_color_coding.rds")
phenotypes <- 
  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
sites <- read_rds("../data/sites.rds")
```
```{r}
phe_df <- phenotypes %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
  select(-PLOT_GL, -WHERE, -YEAR) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)

>>>>

#
```{r}
metadata <- read_rds("../data/metadata.rds")
subpops <- read_rds("../data/subpop_color_coding.rds")
phenotypes <- 
  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
sites <- read_rds("../data/sites.rds")
```

```{r}
phe_df <- phenotypes %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
  select(-PLOT_GL, -WHERE, -YEAR) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean)

### 2020-12-03 R/Pre_manuscript/Flowering_time_environmental_index_analysis.Rmd
Move R scripts not used in current manuscript and rename those that are

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(corrplot)
library(zoo)
```

@@

metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")
phe_all <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_long_non_Blackwell_replants_pre_and_post_2019_replant.rds")
sites <- read_rds("~/Github/pvdiv-phenotypes/data/sites.rds")
fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
```

@@

Make GDD & PTT using a range of base temperatures
```{r}

Tbase <- seq(0, 6)
Topt <- seq(8,30)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(corrplot)
library(zoo)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

@@

metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")
phe_all <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_long_non_Blackwell_replants_pre_and_post_2019_replant.rds")
sites <- read_rds("~/Github/pvdiv-phenotypes/data/sites.rds")
fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
metadata$ECOTYPE <- factor(metadata$ECOTYPE, levels = c("Upland", "Coastal", "Lowland", "Unknown"))
metadata$GENETIC_SUBPOP_50PER <- factor(metadata$GENETIC_SUBPOP_50PER, levels = c("Midwest", "Atlantic", "Gulf", "Admixed", "Unknown"))
```

@@

# Make GDD & PTT using a range of base temperatures
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

@@

# Make GDD & PTT using a range of base temperatures

And starting every individual's GDD and PTT from its own GR50 time
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = ceiling(max(FL50)))

tmp1 <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
  left_join(gwas_d2f)

    tmp <- tmp1 %>%
      group_by(SITE, PLANT_ID, Day_of_year) %>%
      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
      mutate(TMEAN = (TMAX_SM + TMIN_SM) / 2,
             GDD = ifelse(TMEAN >= j, TMEAN - j, 0),
             GDD = ifelse(TMEAN >= k, k - j, GDD),
             PT = GDD * DYLN) %>%
      ungroup %>%
      group_by(SITE, PLANT_ID) %>%
      summarize_at(vars(GDD, PT), funs(sum))
    colnames(tmp)[3:4] = c(paste0('GDD_', j, "_", k), paste0('PT_', j, "_", k))
    GDD_PT = c(GDD_PT, tmp)
    }
}

  GDD_PT <- GDD_PT %>%
    select(SITE, PLANT_ID, starts_with("GDD"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
    ungroup()

# corrplot::corrplot.mixed(M[1:15,1:15])

ggplot(data = M, aes(x = D2F, y = PT_8_33)) + 
  geom_point()
M2 <- enframe(M[1,], name = "PT", value = "corr") %>%
  filter(PT != "D2F") %>%
  separate(col = PT, into = c("Var", "Tbase", "Topt")) %>%
  dplyr::select(-Var) %>%
  pivot_wider(names_from = Topt, values_from = corr, names_prefix = "Topt_") %>%
  mutate(Tbase = paste0("Tbase_", Tbase))
M3 <- M2 %>%
  column_to_rownames(var = "Tbase") %>%
  as.matrix(.)
corrplot::corrplot.mixed(M3)
```
```{r}
write_rds(gwas_avegdd, path = "../analysis/environmental-index/GDD_and_PTT_GWAS.rds")
```

# Do analysis just using daylength, not temperature

And starting every individual's daylength received from its own GR50 time
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = ceiling(max(FL50)))

tmp1 <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
  left_join(gwas_d2f)

    tmp <- tmp1 %>%
      group_by(SITE, PLANT_ID, Day_of_year) %>%
      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
      ungroup %>%
      group_by(SITE, PLANT_ID) %>%
      summarize_at(vars(DYLN), funs(sum))
    GDD_PT = c(GDD_PT, tmp)

  GDD_PT <- GDD_PT %>%
    select(SITE, PLANT_ID, starts_with("DYLN"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
    ungroup()

M = cor(gwas_avegdd %>% select(D2F, starts_with('DYLN')) %>% filter(complete.cases(.)))

 corrplot::corrplot.mixed(M)

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PLANT_ID %in% c("AP13", "J001.B", "J005.B", "J006.B", "J016.B", "J024.A", "J026.B", "J350.C", "J499.B", "J517.B")) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = PLANT_ID)) + 
  geom_smooth(aes(color = PLANT_ID), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE)

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN.8)) + 
  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN.8)) + 
  geom_point(aes(color = ECOTYPE)) + 
  geom_smooth(aes(color = ECOTYPE), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")
```

```{r}
gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
  filter(!(ECOTYPE %in% c("Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = TRUE, method = "gam",
              formula = y ~ poly(x,2)) +
  scale_color_viridis(discrete = TRUE, end = 0.8) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")

save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Genetic_subpop.png", plot = last_plot())

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
  filter(!(ECOTYPE %in% c("Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = ECOTYPE)) + 
  geom_smooth(aes(color = ECOTYPE), se = TRUE, method = "gam", 
              formula = y ~ poly(x,2)) +
  scale_color_viridis(discrete = TRUE, end = 0.8) + 
  ylab("One site-year of cumulative daylength") + 
  xlab("Days to flowering (FL50 - GR50)") + 
  theme(legend.position = "right")

save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Ecotype.png", plot = last_plot())

```

### 2020-12-03 R/Pre_manuscript/GRM_kinship_for_fourway.Rmd
Move R scripts not used in current manuscript and rename those that are

##

@@

geno_df <- cross$geno[[1]]$data

for(i in 2:18){
geno_df <- geno_df %>%
  cbind(cross$geno[[i]]$data)
}

@@

snps <- as.matrix(geno_df2)
# snps[1:20,1000:1005]
```

@@

mash_plot_pairwise_sharing(corrmatrix = K[1:30,1:30])

write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_5250g_fourway_cross.rds")
```

>>>>

## 
```{r}
write_rds(cross, "../data/fourway_cross_data.rds", compress = "gz")
```

@@

geno_df <- cross$geno[[1]]$data
as_tibble(cross$pheno)
for(i in 2:18){
geno_df <- geno_df %>%
  cbind(cross$geno[[i]]$data)
}

@@

snps <- as.matrix(geno_df2)
# snps[1:20,1000:1005]
snps1 <- snps[1:750,]
snps <- snps1
```

@@

mash_plot_pairwise_sharing(corrmatrix = K[1:20,1:20])

write_rds(K, path = "~/Github/pvdiv-phenology-gxe/data/Kinship_van_Raden_750g_fourway_cross_withInbred.rds")
```

```{r}
mash_plot_pairwise_sharing(corrmatrix = K[1:20,1:20])

```

### 2020-12-03 R/Pre_manuscript/Make_mash_U_hyp.Rmd
Move R scripts not used in current manuscript and rename those that are

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
suffix = "gulf_uhyp1"
mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 8)

@@

For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.
```{r}
phe_gulf <- phenotypes$Gulf %>%
  select(PLANT_ID, ends_with("FL50"))

@@

>>>>

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
suffix = "gulf_uhyp1"
mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 8, saveoutput = TRUE)

@@

For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.

No, this doesn't work, because then there are no "0" effects in the random effects. Humm.
```{r}
phe_gulf <- phenotypes$Gulf %>%
  select(PLANT_ID, ends_with("FL50"))

@@

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
numSNPs <- 30000
suffix = "gulf_solstice"
Ugulfhyp <- read_rds(path = file.path(path, "U_hyp_gulf_1phe.rds"))
U_h2 <- list(Udist = Ugulfhyp$U_dist, UGDD = Ugulfhyp$U_CGDD_12C)

#U_ed <- file.path(path, "Mash_data_driven_covariances_30000_gulf_uhyp1.rds")

m <- mash_standard_run(path = path, numSNPs = numSNPs, suffix = suffix,
                       saveoutput = TRUE)
```

Maybe... I could scale everything by the maximum value across the whole dataset. This would be a different type of scaling that could not handle multiple different phenotypes. But, for the same phenotype, it would 1) allow covariance matrices to be scaled from -1 to 1; 2) allow us to still visualize the fact that effect sizes are different at different sites - that there is variance-changing GxE. I think. Hopefully.

I should think a little more on how mash is dealing with this... and look into the mash code...

## Plot eigenvectors to capture predominant patterns in covariance matrices.
```{r}
sum(m$fitted_g$pi)
m$fitted_g$pi[which(m$fitted_g$pi > 0.005)]

v <- m$fitted_g$Ulist$ED_PCA_1
b1=svd(v)
for(g in 1){
barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
}

b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix.

m$fitted_g$grid
```
Plot eigenvectors capturing predominant patterns

The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
```
col = as.character(color.gtex[,2])
for(g in 1:3){
  v=svd(covmat[[k]])$v[h,];d=svd(covmat[[k]])$d
  par(mar=c(8,4.1,4.1,2.1))
  barplot(v[,g]/v[which.max(abs(v[,g])),g],las=2,
          main=paste("Eigenvector",g,"of Uk",k),
          cex.names = 0.5,col=col[h],names=names[h])
}
```
# Matrix math with V_j

Each covariance matrix Vj is specified as Vj = SjCSj, where C is a correlation matrix that accounts for correlations among the measurements in the R conditions, and Sj is the R x R diagonal matrix with diagonal elements s^j1, … , s^jR.  If the measurements in the R conditions are independent, one would set C = IR, the R x R identity matrix, and therefore Vj = Sj2.

Interestingly, they assume row-wise normal distribution of effect sizes, so a SNP has effects taken from a normal distribution. They assume b^j is normally distributed about bj with variance-covariance matrix Vj. The more highly correlated two conditions are, the 
```{r}
Sj <- matrix(data = c(3, 0 , 0, 2), nrow = 2, ncol = 2)
Clo <- matrix(data = c(1, 0.1, 0.1, 1), nrow = 2, ncol = 2)
isSymmetric(Clo)
Cid <- matrix(data = c(1, 0, 0, 1), nrow = 2, ncol = 2)
Chi <- matrix(data = c(1, 1, 1, 1), nrow = 2, ncol = 2)

Sj %*% Cid %*% Sj
Sj %*% Clo %*% Sj
Sj %*% Chi %*% Sj
?cov2cor

```{r}
dataf = list(Bhat = B_hat_random_df_30000topSNPs_gulf_uhyp1, Shat = S_hat_random_df_30000topSNPs_gulf_uhyp1)
estimate_null_correlation_simple = function(dataf, z_thresh=2, est_cor = TRUE){
  z = dataf$Bhat/dataf$Shat
  max_absz = apply(abs(z),1, max)
  nullish = which(max_absz < z_thresh)
  if(length(nullish)<n_conditions(dataf)){
    stop("not enough null data to estimate null correlation")
  }
  nullish_z = z[nullish,]
  Vhat = cov(nullish_z)
  if(est_cor){
    Vhat = cov2cor(Vhat)
  }
  return(Vhat)
}
```
